---
title: "Canadian Doctors Iterations 3"
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)  # Load the scales package for percentage formatting
library(knitr)

if(Sys.info()['user']=='gr2757') {
  
  Project <- "/Users/gr2757/Dropbox/CPSO/output"
}

if(Sys.info()['user']=='T-800 Model 101') {
  
  Project <- "D:/Dropbox/CPSO/output"
}
if(Sys.info()['user']=='wojtek') {
#    project<-"/home/wojtek/Dropbox/MyPapers/Canada-Incorporation/project_incorp/incorp_results/Results/20250206"
}


all_doctors_path = file.path(Project, "all_doctors_data_three.csv")

all_doctors_raw = read.csv(all_doctors_path, stringsAsFactors = FALSE)
all_doctors_raw[all_doctors_raw == ""] <- NA

#setwd("D:/Dropbox/CPSO/graphing")

```

```{r}
#keeping these columns

#print(colnames(all_doctors_raw))
all_doctors = subset(all_doctors_raw, select = c("cpso", "name", "active_member", "corporation_date_1", "expired_date", "left_corporation_1", "expired_corporation_date_1",  "specialties_1", "specialties_2", "specialties_3", "specialties_4", "educ", "educ_year"))

all_doctors = all_doctors %>%
  rename("corporation_date" = "corporation_date_1") %>%
  rename("expired_corporation_date" = "expired_corporation_date_1") %>%
  rename("specialties" = "specialties_1") %>%
  rename("left_corporation" = "left_corporation_1")

print(colnames(all_doctors))

#make blank into NA
all_doctors[all_doctors == ""] <- NA

print(sum(is.na(all_doctors$active_member)))
print(sum(is.na(all_doctors$expired_date)))



#create a count column
all_doctors = all_doctors %>%
  mutate(specialties_count = rowSums(!is.na(across(starts_with("specialties")))))

```

```{r}
#plot of specialties count
ggplot(all_doctors, aes(x = specialties_count)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.5) +  # Add labels above bars
  labs(title = "Number of Specialties per Doctor", x = "Number of Specialties", y = "Count of Doctors") +
  theme_minimal()
```
Zero is NA effectively. 

```{r}

# adding is_incorp column
all_doctors <- all_doctors %>%
  mutate(is_incorp = ifelse(!is.na(corporation_date) | !is.na(expired_corporation_date), 1, 0))

#only looking at doctors with listed specialties
with_specialties <- all_doctors %>%
  filter(specialties_count > 0) %>%
  mutate(specialties = gsub("and", "&", specialties, ignore.case = TRUE))

incorp_counts <- with_specialties %>%
  filter(is_incorp == 1) %>%
  count(specialties, name = "count") %>%
  arrange(desc(count)) %>%
  slice_max(count, n = 30)  # Keep top 30 specialties


ggplot(incorp_counts, aes(x = reorder(specialties, count), y = count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = count), hjust = -.03, size = 3, nudge_y = .5) +  # Adjust text size
  coord_flip() +
  labs(title = "Top 30 Specialties with Incorporated Doctors",
       x = "Specialty",
       y = "Number of Incorporated Doctors") +
  theme_minimal() +
  theme(axis.text.y = element_text(margin = margin(r = -15)))  # Adjust right margin for y-axis labels

```
Family Medicine count is cut off, the value is 8,428

```{r}
# all doctors with specialties

all_specialties_count <- all_doctors %>%
  filter(specialties_count > 0) %>%
  mutate(specialties = gsub("and", "&", specialties, ignore.case = TRUE)) %>%
  count(specialties, name = "count") %>%
  arrange(desc(count)) %>%
  slice_max(count, n = 30)  # Keep top 30 specialties

ggplot(all_specialties_count, aes(x = reorder(specialties, count), y = count)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = count), hjust = -.03, size = 3, nudge_y = .5) +  # Adjust text size
  coord_flip() +
  labs(title = "Top 30 Specialties (All Doctors)",
       x = "Specialty",
       y = "Number of Doctors") +
  theme_minimal() +
  theme(axis.text.y = element_text(margin = margin(r = -15)))  # Adjust right margin for y-axis labels

```
The count for Family Medicine is 16,699. It seems that the top 30 specialties look pretty similar, incorporated or not.

```{r}
incorp_proportions <- with_specialties %>%
  group_by(specialties) %>%
  summarise(count_incorp = sum(is_incorp == 1), count_total = n()) %>%
  mutate(proportion = count_incorp / count_total) %>%
  slice_max(proportion, n = 30)  # Keep top 30 specialties by proportion

# Create the bar plot for proportions
ggplot(incorp_proportions, aes(x = reorder(specialties, proportion), y = proportion)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = paste0(round(proportion * 100), "%")), hjust = -.05, size = 3) +  # Add rounded percentage labels
  coord_flip() +
  labs(title = "Top 30 Specialties by Proportion \n of Incorporated Doctors",
       x = "Specialty",
       y = "Proportion of Incorporated Doctors") +
  theme_minimal() +
  theme(axis.text.y = element_text(margin = margin(r = 10)))
```

The top 5 specialties by proportion is due to how few doctors have these specialties. Addiction medicine, Pediatric Gastroenterology, and Cardiothoracic Surgery have 1, 1, 3 doctors practicing those specialties; they just all happen to be incorporated. Vascular surgery has 57 doctors and 28 of the are incorporated. Lastly, Family medicine has a total of 16,699 doctors, and 8,428 are incorporated. 



```{r}
#table
incorp_proportions_table = incorp_proportions %>%
  arrange(desc(count_total)) %>%
  top_n(30) %>%
  rename("num_doctors" = "count_incorp") %>%
  rename("incorporation_rate" = "proportion") %>%
  select(-c(count_total))

incorp_proportions_table$incorporation_rate = percent(incorp_proportions_table$incorporation_rate, accuracy = .1)

kable(incorp_proportions_table)

```


```{r}
# Create a new column for the year of incorporation
expanded_doctors = subset(all_doctors, select = -c(educ, specialties, specialties_2, specialties_3, specialties_4, specialties_count))

# turning dates into just year
expanded_doctors <- expanded_doctors %>%
  mutate(incorp_year = as.numeric(format(as.Date(corporation_date, format = "%Y-%m-%d"), "%Y")))

expanded_doctors <- expanded_doctors %>%
  mutate(expired_year = as.numeric(format(as.Date(expired_date, format = "%Y-%m-%d"), "%Y")))

expanded_doctors <- expanded_doctors %>%
  mutate(expired_corp_year = as.numeric(format(as.Date(expired_corporation_date, format = "%Y-%m-%d"), "%Y")))

expanded_doctors <- expanded_doctors %>%
  mutate(active_year = as.numeric(format(as.Date(active_member, format = "%Y-%m-%d"), "%Y")))


expanded_doctors <- expanded_doctors %>%
  mutate(left_corp_year = as.numeric(format(as.Date(left_corporation, format = "%Y-%m-%d"), "%Y")))

#drop true NAs
print(paste("No good year to choose for this many rows: ", sum(is.na(all_doctors$educ_year))))
expanded_doctors = expanded_doctors %>%
  drop_na(educ_year)

#the next best thing to when a doctor became active is when they graduated med school
expanded_doctors = expanded_doctors %>%
  mutate(active_year = coalesce(active_year, educ_year)) #uses educ_year if active_year is NA

#drop the date columns
expanded_doctors = subset(expanded_doctors, select = -c(corporation_date, expired_corporation_date, expired_date, active_member, left_corporation))


#if there is value for active year, but not expired year, set expired year to 2025 (they're still active to this day)
expanded_doctors <- expanded_doctors %>%
  mutate(expired_year = ifelse(!is.na(active_year) & is.na(expired_year), 2025, expired_year)) #if there is active year, set expired year to 2025

#if there is a value for incorp year, but not expired corp year, set expired corp year to 2025 (they're still active to this day)
expanded_doctors <- expanded_doctors %>%
  mutate(expired_corp_year = ifelse(!is.na(incorp_year) & is.na(expired_corp_year), 2025, expired_corp_year))

print(sum(!is.na(expanded_doctors$incorp_year) & !is.na(expanded_doctors$expired_corp_year)))

# Expand the dataset
expanded_doctors <- expanded_doctors %>%
  rowwise() %>%
  mutate(years = list(seq(active_year, as.integer(expired_year)))) %>%  # Create a sequence of years
  unnest(years)

expanded_doctors <- expanded_doctors %>%
  drop_na(expired_year, active_year) %>%  # Ensure necessary data is present
  mutate(is_active_status = ifelse(years >= active_year & years <= expired_year, 1, 0)) %>% # Create a binary variable for active status (all 1)
  mutate(is_incorp_status = ifelse(
    is.na(incorp_year) | is.na(expired_corp_year), 0,  # If either is NA, set to 0
    ifelse(years >= incorp_year & years <= expired_corp_year, 1, 0)  # Otherwise, apply the range condition
  ))

```


```{r}
incorp_over_time <- expanded_doctors %>%
  filter(is_incorp == 1) %>%  # Keep only incorporated doctors
  drop_na(active_year)  # Ensure necessary data is present

temp_df = expanded_doctors %>%
  filter(expired_corp_year != 2025)


incorp_over_time_counts <- expanded_doctors %>%
  group_by(years) %>%  # Group by year
  summarise(total_incorp = sum(is_incorp_status, na.rm = TRUE)) %>%
  arrange(years)  # Arrange by year for plotting



# Filter the dataset to start from the selected year
incorp_over_time_counts <- incorp_over_time_counts %>%
  filter(years >= 2000 & years <= 2024)


# Create the plot
ggplot(incorp_over_time_counts, aes(x = years, y = total_incorp)) +
  geom_line(color = "blue", linewidth = 1) +  # Line plot
  geom_point(color = "blue") +  # Add points to the line
  labs(title = "Number of Active Incorporated Doctors Over Time",
       x = "Year",
     y = "Number of Incorporated Doctors") +
  theme_minimal()
```

I fixed the start of the graph to be 5 years before the first year of incorporation, the first incorporation was 2002 with one, and 2003 had 239 new incorporations.

This graph is missing any doctors that were incorporated but are no longer active. So there is never any decrease in the number of incorporated doctors. So if a doctor was incorporated in 2005 and retired in 2020, they would have never shown up in the graph at all.

```{r}
# incorp over time as a proportion of all doctors (only active doctors per year)
incorp_over_time_prop <- expanded_doctors %>%
  drop_na(expired_year, active_year) %>%  # Ensure necessary data is present
  mutate(is_incorp_status = ifelse(years >= incorp_year & years <= expired_corp_year, 1, 0))

incorp_over_time_prop <- incorp_over_time_prop %>%
  group_by(years) %>%  # Group by year
  summarise(
    total_incorp = sum(is_incorp_status, na.rm = TRUE),  # Count incorporated doctors
    total_active = sum(is_active_status, na.rm = TRUE),  # Count active doctors
    prop_incorp = total_incorp / total_active  # Compute proportion
  ) %>%
  arrange(years)  # Arrange by year for plotting
  
# Filter the dataset to start from the selected year
incorp_over_time_prop <- incorp_over_time_prop %>%
  filter(years >= 2000 & years <= 2024)

# Create the plot
ggplot(incorp_over_time_prop, aes(x = years, y = prop_incorp)) +
  geom_line(color = "blue", linewidth = 1) +  # Line plot
  geom_point(color = "blue") +  # Add points to the line
  labs(title = "The Proportion of All Doctors That are Incorporated",
       x = "Year",
       y = "Percentage of Incorporated Doctors") +
  scale_y_continuous(labels = percent) +  # Convert y-axis labels to percentages
  theme_minimal()

```
The  doctors that are listed as inactive are missing data for when they incorporated, but do include when they corporation status expired. Doctors will either have a listed incorporation date or an expiration date for their corporation status, but not both.

The problem persists of missing historic data with inactive doctors and their missing data. 


```{r}
# going by cohort

expanded_doctors_cohort <- expanded_doctors %>%
  drop_na(educ_year) %>%  # Ensure necessary data is present
  mutate(cohort = cut(educ_year, 
                      breaks = seq(1950, 2030, by = 10),  # 10-year bins
                      labels = paste0(seq(1950, 2020, by = 10), "-", seq(1959, 2029, by = 10)),
                      include.lowest = TRUE))


active_doctors_cohort <- expanded_doctors_cohort %>%
  drop_na(cohort) %>%
  group_by(cohort, years) %>%
  summarise(total_active = sum(is_active_status, na.rm = TRUE)) %>%
  ungroup()

# Plot total active doctors by cohort over time
ggplot(active_doctors_cohort, aes(x = years, y = total_active, color = as.factor(cohort))) +
  geom_line(linewidth = 1) +  # Line plot
  geom_point() +  # Add points to the line
  labs(
    title = "Total Active Doctors per Cohort Over Time",
    x = "Year",
    y = "Total Active Doctors",
    color = "Graduation Cohort"
  ) +
  theme_minimal()

```
There is data for when doctors became active, either explicitly through doctors that are active and have a listed year for when they became active or implicitly by using the year they graduated med school for inactive doctors. This is what I would expect the incorporation by cohorts to look like if there were solid start and end dates for incorporation.  


```{r}
expanded_doctors_cohort_counts  <- expanded_doctors_cohort %>%
  drop_na(cohort) %>%
  filter(years >= 2000 & years <= 2024) %>% 
  group_by(cohort, years) %>%
  summarise(total_incorp = sum(is_incorp_status, na.rm = TRUE)) %>%
  ungroup()

# Plot total incorporated doctors by cohort over time
ggplot(expanded_doctors_cohort_counts, aes(x = years, y = total_incorp, color = as.factor(cohort))) +
  geom_line(linewidth = 1) +  # Line plot
  geom_point() +  # Add points to the line
  labs(
    title = "Total Incorporated Doctors per Cohort Over Time",
    x = "Year",
    y = "Total Incorporated Doctors",
    color = "Graduation Cohort"
  ) +
  theme_minimal()


```
I set the graph to start in 2000, 2 years before the first incorporation. Some doctors have an incorporation date before they graduated med school, this is likely due to them joining an already incorporated practice.

```{r}

expanded_doctors_cohort_prop <- expanded_doctors_cohort %>%
  group_by(cohort, years) %>%
  summarise(
    total_incorp = sum(is_incorp_status, na.rm = TRUE),  # Count incorporated doctors
    total_doctors = sum(is_active_status == 1, na.rm = TRUE),  # Count only active doctors
    prop_incorp = total_incorp / total_doctors  # Compute proportion
  ) %>%
  ungroup()

expanded_doctors_cohort_prop <- expanded_doctors_cohort_prop %>%
  filter(years >= 2000 & years <= 2024) %>%
  drop_na(cohort)  # Ensure necessary data is present


ggplot(expanded_doctors_cohort_prop, aes(x = years, y = prop_incorp, color = as.factor(cohort))) +
  geom_line(linewidth = 1) +  # Line plot
  geom_point() +  # Add points to the line
  labs(
    title = "Proportion of Incorporated Doctors per Active Cohort Over Time",
    x = "Year",
    y = "% of Incorporated Doctors",
    color = "Graduation Cohort"
  ) +
  scale_y_continuous(labels = percent) +  # Format y-axis as percentages
  theme_minimal()

```
This is not informative since there is lacking of any expiration of incorporation.


```{r}
doctors_chort_five <- expanded_doctors %>%
  mutate(cohort = cut(educ_year, 
                      breaks = seq(1950, 2030, by = 5),  
                      labels = paste0(seq(1950, 2025, by = 5), "-", seq(1954, 2029, by = 5)),
                      include.lowest = TRUE))

doctors_chort_five_counts  <- doctors_chort_five %>%
  drop_na(cohort) %>%
  filter(years >= 2000 & years <= 2024) %>% 
  group_by(cohort, years) %>%
  summarise(total_incorp = sum(is_incorp_status, na.rm = TRUE)) %>%
  ungroup()

# Plot total incorporated doctors by cohort over time
ggplot(doctors_chort_five_counts, aes(x = years, y = total_incorp, color = as.factor(cohort))) +
  geom_line(linewidth = 1) +  # Line plot
  geom_point() +  # Add points to the line
  labs(
    title = "Total Incorporated Doctors per Cohort Over Time",
    x = "Year",
    y = "Total Incorporated Doctors",
    color = "Graduation Cohort"
  ) +
  theme_minimal()



```


```{r}
# Step 3: Summarize incorporation over time for each cohort
doctors_chort_five_prop <- doctors_chort_five %>%
  group_by(cohort, years) %>%
  summarise(
    total_incorp = sum(is_incorp_status, na.rm = TRUE),  # Count incorporated doctors
    total_doctors = sum(is_active_status == 1, na.rm = TRUE),  # Count only active doctors
    prop_incorp = total_incorp / total_doctors  # Compute proportion
  ) %>%
  ungroup()
  
doctors_chort_five_prop <- doctors_chort_five_prop %>%
  filter(years >= 2000 & years <= 2024) %>%
  drop_na(cohort)  # Ensure necessary data is present


# Step 5: Plot cumulative incorporations over time by cohort
ggplot(doctors_chort_five_prop, aes(x = years, y = prop_incorp, color = cohort, group = cohort)) +
  geom_line(linewidth = 1) +  # Line plot
  geom_point() +  # Add points to the line
  labs(
    title = "Proportion of Incorporated Doctors per Active Cohort Over Time",
    x = "Year",
    y = "% of Incorporated Doctors",
    color = "Graduation Cohort"
  ) +
  scale_y_continuous(labels = percent) +  # Format y-axis as percentages
  theme_minimal()

```

```{r}
incorp_doctors <- all_doctors %>%
  mutate(incorp_year = as.numeric(format(as.Date(corporation_date, format = "%Y-%m-%d"), "%Y"))) %>%
  filter(is_incorp == 1) %>%
  mutate(incorp_after_med_school = incorp_year - educ_year) %>%
  drop_na(educ_year) %>%  # Ensure necessary data is present
  mutate(cohort = cut(educ_year, 
                      breaks = seq(1950, 2030, by = 10),  # 10-year bins
                      labels = paste0(seq(1950, 2020, by = 10), "-", seq(1959, 2029, by = 10)),
                      include.lowest = TRUE))


avg_incorp_by_cohort <- incorp_doctors %>%
  group_by(cohort) %>%
  summarise(avg_years_to_incorp = mean(incorp_after_med_school, na.rm = TRUE)) %>% 
  drop_na(avg_years_to_incorp) %>%  # Ensure necessary data is present
  ungroup()

# Step 3: Create the bar chart
ggplot(avg_incorp_by_cohort, aes(x = cohort, y = avg_years_to_incorp, fill = cohort)) +
  geom_col() +  # Creates a bar chart
  geom_text(aes(label = round(avg_years_to_incorp, 1)), vjust = -0.5) +  # Adds labels above bars
  labs(title = "Average Years to Incorporation After Med School by 10 Year Cohort",
       x = "Cohort",
       y = "Average Years to Incorporation") +
  theme_minimal() +
  theme(legend.position = "none")  # Hide legend since fill is same as x-axis

```
Some of the averages include negative values pulling down the average. This is due to new doctors joining incorporated practice that where incorporated before they graduated med school. There are negative values for only 48 of the doctors. 



```{r}
#add column for whether they're active or inactive

all_doctors_raw <- all_doctors_raw %>%
  mutate(is_active = ifelse(is.na(expired_date), 1, 0))

active_doctors <- all_doctors_raw %>%
  mutate(active_year = as.numeric(format(as.Date(active_member, format = "%Y-%m-%d"), "%Y")))

active_doctors$gap = active_doctors$active_year - active_doctors$educ_year


#print(median(active_doctors$gap, na.rm = TRUE))
#print(mean(active_doctors$gap, na.rm = TRUE))

gap_df <- active_doctors %>%
  filter(is_active == 1) %>%
  drop_na(gap) %>%
  group_by(gap) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  mutate(prop = round(count / sum(count), 2)) %>%
  arrange(desc(prop)) %>%
  slice(1:11) %>%
  arrange(gap)

kable(gap_df)
```

The average amount of time for a doctor to become active after med school is 5.4 years, and the median is 2 years. 

```{r}

both_doctors <- all_doctors_raw %>%
  rename("corporation_date" = "corporation_date_1") %>%
  rename("expired_corporation_date" = "expired_corporation_date_1") %>%
  rename("left_corporation" = "left_corporation_1") %>%
  rename("corporation_name" = "corporation_name_1") %>%
  mutate(expired_year = as.numeric(format(as.Date(expired_date, format = "%Y-%m-%d"), "%Y"))) %>%
  mutate(active_year = as.numeric(format(as.Date(active_member, format = "%Y-%m-%d"), "%Y"))) %>%
  mutate(left_corp_year = as.numeric(format(as.Date(left_corporation, format = "%Y-%m-%d"), "%Y"))) %>%
  mutate(incorp_year = as.numeric(format(as.Date(corporation_date, format = "%Y-%m-%d"), "%Y"))) %>%
  mutate(expired_corp_year = as.numeric(format(as.Date(expired_corporation_date, format = "%Y-%m-%d"), "%Y")))

#removing the date columns that have a year column
both_doctors <- both_doctors %>%
  select(-c(expired_date, corporation_date, active_member, left_corporation, corporation_date, expired_corporation_date))

#the next best thing to when a doctor became active is when they graduated med school
both_doctors = both_doctors %>%
  mutate(active_year = coalesce(active_year, educ_year)) #uses educ_year if active_year is NA

#add a binary for has corporation and counts for other columns 
both_doctors <- both_doctors %>%
  mutate(has_corp = ifelse(!is.na(corporation_name) | !is.na(expired_corp_year) | !is.na(left_corp_year), 1, 0)) %>% #has ever incorporated
  mutate(corp_count = rowSums(!is.na(across(starts_with("corp_address"))))) %>%
  mutate(shareholder_count = rowSums(!is.na(across(starts_with("corporate_shareholders"))))) %>%
  mutate(hospital_count = rowSums(!is.na(across(starts_with("hospital_privileges")))))


#if there is value for active year, but not expired year, set expired year to 2025 (they're still active to this day)
both_doctors <- both_doctors %>%
  mutate(expired_year = ifelse(!is.na(active_year) & is.na(expired_year), 2025, expired_year)) #if there is active year, set expired year to 2025

#if there is a value for incorp year, but not expired corp year, set expired corp year to 2025 (they're still active to this day)
both_doctors <- both_doctors %>%
  mutate(expired_corp_year = ifelse(!is.na(incorp_year) & is.na(expired_corp_year), 2025, expired_corp_year))

both_doctors <- both_doctors %>%
  mutate(expired_corp_year = ifelse(!is.na(incorp_year) & is.na(left_corp_year), 2025, left_corp_year))


#dropping unused columns after they were added to counts
both_doctors <- both_doctors %>%
  select(-c(starts_with("corporate_shareholders"), starts_with("corporate_shareholders"), starts_with("corp_address"),    starts_with("hospital_privileges"), starts_with("specialties")))

#print(sum(!is.na(both_doctors$left_corporation)))



# Expand the dataset
both_doctors <- both_doctors %>%
  drop_na(expired_year, active_year) %>%  # Ensure necessary data is present
  rowwise() %>%
  mutate(years = list(seq(active_year, as.integer(expired_year)))) %>%  # Create a sequence of years, from when they start to experation
  unnest(years)

both_doctors <- both_doctors %>%
  mutate(is_active_status = ifelse(years >= active_year & years <= expired_year, 1, 0)) %>% # Create a binary variable for active status (all 1)
  mutate(is_incorp_status = ifelse(
    is.na(incorp_year) | is.na(left_corp_year), 0,  # If either is NA, set to 0
    ifelse(years >= incorp_year & years <= left_corp_year, 1, 0)  # Otherwise, apply the range condition
  ))


```



```{r}
#inactive doctors only

inactive_doctors <- both_doctors %>%
  filter(is_active == 0) %>%
  drop_na(educ_year) %>%  # Ensure necessary data is present
  mutate(cohort = cut(educ_year, 
                      breaks = seq(1950, 2030, by = 10),  # 10-year bins
                      labels = paste0(seq(1950, 2020, by = 10), "-", seq(1959, 2029, by = 10)),
                      include.lowest = TRUE))

inactive_doctors_counts  <- inactive_doctors %>%
  drop_na(cohort) %>%
  filter(years >= 1980 & years <= 2024) %>% 
  group_by(cohort, years) %>%
  summarise(total_incorp = sum(is_active_status, na.rm = TRUE)) %>%
  ungroup()

# Plot total incorporated doctors by cohort over time
ggplot(inactive_doctors_counts, aes(x = years, y = total_incorp, color = as.factor(cohort))) +
  geom_line(linewidth = 1) +  # Line plot
  geom_point() +  # Add points to the line
  labs(
    title = "Number of Doctors per Cohort Over Time (Inactive Doctors)",
    x = "Year",
    y = "Total Inactive Doctors",
    color = "Graduation Cohort"
  ) +
  theme_minimal()

```

This is a graph showing the history of doctors that are as of now are inactive. 



```{r}

##add cohort
both_doctors_prop <- both_doctors %>%
  mutate(cohort = cut(educ_year, 
                      breaks = seq(1950, 2030, by = 10),  # 10-year bins
                      labels = paste0(seq(1950, 2020, by = 10), "-", seq(1959, 2029, by = 10)),
                      include.lowest = TRUE))



#inactive doctors as a proportion of total doctors (inactive + active)
both_doctors_prop <- both_doctors_prop %>%
  drop_na(cohort) %>%  # Ensure necessary data is present
  group_by(cohort, years) %>%
  summarise(
    inactive_doctors_total = sum(is_active == 0, na.rm = TRUE),  # Count inactive doctors
    active_total_cohort = sum(is_active_status == 1, na.rm = TRUE)  # Active within cohort
  ) %>%
  ungroup() %>%
  group_by(years) %>%  # Now group only by years
  mutate(active_total_doctors = sum(active_total_cohort, na.rm = TRUE)) %>%  # Sum across cohorts
  ungroup() %>%
  mutate(inactive_prop = inactive_doctors_total / active_total_doctors) %>%  # Compute proportion
  filter(years >= 1980 & years <= 2024)


ggplot(both_doctors_prop, aes(x = years, y = inactive_prop, color = as.factor(cohort))) +
  geom_line(linewidth = 1) +  # Line plot
  geom_point() +  # Add points to the line
  labs(
    title = "Proportion of Inactive Doctors to Total Doctors per Cohort",
    x = "Year",
    y = "Proportion of Inactive Doctors to Total Docotrs",
    color = "Graduation Cohort"
  ) +
  theme_minimal()

```

An inactive doctor is a doctor that is no longer active as of when the data was collected. So this is showing an inactive doctor today, when they were active in the past. All the lines trend to zero due to active doctors will be a larger and large amount the closer you get to the present. 


```{r}
#share of doctors with a corp listed

inactive_doctors_corp <- both_doctors %>%
  filter(is_active == 0) %>% #all inactive doctors only
  drop_na(educ_year) %>%  # Ensure necessary data is present
  mutate(cohort = cut(educ_year, 
                      breaks = seq(1950, 2030, by = 10),  # 10-year bins
                      labels = paste0(seq(1950, 2020, by = 10), "-", seq(1959, 2029, by = 10)),
                      include.lowest = TRUE))

inactive_doctors_corp_prop  <- inactive_doctors_corp %>%
  drop_na(cohort) %>%
  filter(years >= 2000 & years <= 2024) %>%
  group_by(cohort, years) %>%
  summarise(
    total_doctors_cohort = sum(is_active == 0, na.rm = TRUE),  # Count inactive doctors within cohort
    with_corp_count = sum(has_corp == 1, na.rm = TRUE)  # Count corp within cohort
  ) %>%
  ungroup() %>%
  group_by(years) %>%  # Now group only by years
  mutate(inactive_total_doctors = sum(total_doctors_cohort, na.rm = TRUE)) %>%  # Sum across cohorts
  mutate(with_corp_prop = with_corp_count / inactive_total_doctors)  # Compute proportion


# Plot total incorporated doctors by cohort over time
ggplot(inactive_doctors_corp_prop, aes(x = years, y = with_corp_prop, color = as.factor(cohort))) +
  geom_line(linewidth = 1) +  # Line plot
  geom_point() +  # Add points to the line
  labs(
    title = "Proportion of Inactive Doctors with a Corporation",
    x = "Year",
    y = "Proportion",
    color = "Graduation Cohort"
  ) +
  theme_minimal()


```
An inactive doctor counts as having a corporation if they have a corporation name, ever left a corporation or have an expiration date for their corporation. The proportion is how many doctors within a cohort that ever had a corporation divided by all active doctors at the time (who no longer practice today).



```{r}
# number of doctors with the same corporation listed

```


