---
title: "Canadian Doctors Iteration 5"
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

```

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)  # Load the scales package for percentage formatting
library(knitr)
library(stringr)

if(Sys.info()['user']=='gr2757') {
  
  Project <- "/####/Dropbox/CPSO/output"
}



all_doctors_path = file.path(Project, "all_doctors_data_three.csv")

all_doctors_raw = read.csv(all_doctors_path, stringsAsFactors = FALSE)
all_doctors_raw[all_doctors_raw == ""] <- NA

```



```{r}

#make some counts and drop the multiple columns
clean_all_doctors  = all_doctors_raw %>%
  mutate(count_specialties = rowSums(!is.na(across(starts_with("specialties"))))) %>% 
  select(-starts_with("specialties")) %>%
  mutate(count_hospital_privileges = rowSums(!is.na(across(starts_with("hospital_privileges"))))) %>%
  select(-starts_with("hospital_privileges")) %>%
  mutate(count_corp_address = rowSums(!is.na(across(starts_with("corp_address"))))) %>%
  select(-starts_with("corp_address")) %>%
  mutate(count_corporation = rowSums((!is.na(across(starts_with("corporation_name"))))))

#create a variable for whether a doctor is active and has a corporation
clean_all_doctors = clean_all_doctors %>%
  mutate(is_active = ifelse(is.na(expired_date), 1, 0)) %>%
  mutate(has_incorp = ifelse(!is.na(corporation_name_1), 1,0))

#replace doctor names in listed shareholders with just CPSO#
clean_all_doctors <- clean_all_doctors %>%
  mutate(across(starts_with("corporate_shareholder"), 
                ~ as.numeric(gsub("[^0-9]", "", .))))

  
```

```{r}
#making a data set to deal with corporations:
corporation_df = clean_all_doctors %>%
  filter(has_incorp == 1) %>% #only doctor with a corporation
  filter(is_active == 1) %>% #only active doctors
  select(matches("^corporate_shareholders|^corporation_name|^left_corporation"), cpso)


df_long_corporations <- corporation_df %>%
  select(cpso, starts_with("corporation_name"), starts_with("left_corporation")) %>%
  pivot_longer(
    cols = c(starts_with("corporation_name"), starts_with("left_corporation")),
    names_to = c(".value", "corporation_number"),
    names_pattern = "(corporation_name|left_corporation)_(\\d+)"
  ) %>%
  drop_na(corporation_name)%>%
  rename(corporation = corporation_name) %>%
  filter(is.na(left_corporation)) %>% # Drop rows where they left the corporation
  select(-left_corporation) # Drop the left_corporation column

# Reshape shareholder columns into long format, handling multiple shareholder columns per corporation
df_long_shareholders <- corporation_df %>%
  select(cpso, starts_with("corporate_shareholders")) %>%
  pivot_longer(
    cols = starts_with("corporate_shareholders"),
    names_to = c("corporation_number", "shareholder_number"),
    names_pattern = "corporate_shareholders_(\\d+)_(\\d+)",  # Pattern to capture both corporation and shareholder numbers
    values_to = "shareholder"
  ) %>%
  drop_na(shareholder)  # Drop rows where no shareholder is listed


# Combine the reshaped corporation and shareholder data
df_combined <- left_join(df_long_corporations, df_long_shareholders, by = c("corporation_number", "cpso"))

#only keep cpso corporation and shareholder
df_combined <- df_combined %>% 
  select(cpso, corporation, shareholder)

#this so that every doctor that lists a firm is also treated as a shareholder
cpso_shareholder <- df_combined %>%
  select(-shareholder) %>% #drop the current shareholder
  mutate(shareholder = cpso) #use the cpso as a shareholder 

df_combined = rbind(df_combined, cpso_shareholder) #add a lot duplicates

df_combined <- df_combined %>%
  distinct(corporation, shareholder, .keep_all = FALSE) %>% #remove the duplicates, so now firms that had NA but did have a listed doctor as a cpso now carry that value as a shareholder
  drop_na(shareholder)

#this is all to count how many shareholders a corporation has.
corp_count = df_combined %>%
  group_by(corporation) %>% #group by name
  mutate(shareholder_count = n()) #count how many times that name shoes up


```

```{r}
# count how many shareholders per corporation by corporation name and ignoring corporate shareholders.

long_corp_two = corporation_df %>%
  select(cpso, starts_with("corporation_name"), starts_with("left_corporation")) %>%
  pivot_longer(
    cols = c(starts_with("corporation_name"), starts_with("left_corporation")),
    names_to = c(".value", "corporation_number"),
    names_pattern = "(corporation_name|left_corporation)_(\\d+)"
  ) %>%
  drop_na(corporation_name)%>%
  rename(corporation = corporation_name) %>%
  filter(is.na(left_corporation)) %>% # Drop rows where they left the corporation
  select(-c(left_corporation, corporation_number)) # Drop the left_corporation column

#long_shareholders_two = corporation_df %>%
#  select(cpso, starts_with("corporation_name")) %>%
#  pivot_longer(
#    cols = starts_with("corporation_name"),
#    names_to = ("corporation_number"),
#    names_pattern = "corporation_name_(\\d+)",  # Pattern to capture both corporation and shareholder numbers
#  ) %>%
#  rename(corporation = value)

long_shareholders_two <- corporation_df %>%
  select(cpso, starts_with("corporation_name")) %>%
  pivot_longer(
    cols = starts_with("corporation_name"),
    values_to = "corporation_name"   # This will capture the corporation names in a new column
  ) %>%
  filter(!is.na(corporation_name)) %>%
  rename(corporation = corporation_name) %>%# Rename the column to corporation
  select(-name)

combined_df_two = left_join(long_corp_two, long_shareholders_two, by = c("cpso", "corporation"))

combined_df_two = combined_df_two %>% 
  group_by(corporation) %>% #group by name
  mutate(shareholder_count = n()) %>% #count how many times that name shoes up
  distinct(corporation, shareholder_count) #drop duplicates, one observation per corporation.
```

```{r}
#make a table showing this
table_df = combined_df_two %>%
  group_by(shareholder_count) %>%
  summarise(count = n()) %>%
  mutate(percentage =  round(count / nrow(combined_df_two), 4)) %>%
  arrange(shareholder_count)
  
  
  
kable(table_df)

```

```{r}
gender_specialties = all_doctors_raw %>%
  select(gender, specialties_1, cpso, active_member) %>%
  filter(!is.na(specialties_1)) %>%
  filter(!is.na(active_member)) %>%
  filter(!gender == "Unknown") %>%
  mutate(is_women = ifelse(gender == "Female", 1, 0)) %>%
  group_by(specialties_1) %>%
  mutate(specialty_count = n()) %>%
  ungroup()

 
gender_specialties = gender_specialties %>%
  group_by(specialties_1, gender) %>%
  mutate(women_count = n()) %>%
  ungroup() %>%
  filter(gender == "Female") %>%
  distinct(specialties_1, .keep_all = TRUE) %>%
  mutate(percentage = round(women_count / specialty_count * 100, 2)) %>%
  select(c(specialties_1, women_count, specialty_count, percentage)) %>%
  arrange(desc(percentage))



kable(gender_specialties)



```


```{r}
#gender cohort
gender_cohort <- all_doctors_raw %>%
  select(cpso, educ_year, gender, active_member) %>%
  filter(!is.na(active_member)) %>%
  drop_na(educ_year) %>%  # Ensure necessary data is present
  mutate(cohort = cut(educ_year, 
                      breaks = seq(1950, 2030, by = 5),  
                      labels = paste0(seq(1950, 2025, by = 5), "-", seq(1954, 2029, by = 5)),
                      include.lowest = TRUE)) %>%
  drop_na(cohort) %>%
  select(-c(educ_year, active_member)) %>%
  group_by(cohort) %>%
  mutate(cohort_count = n()) %>%
  ungroup()

gender_cohort = gender_cohort %>%
  group_by(cohort, gender) %>%
  mutate(women_cohort_count = n()) %>%
  ungroup() %>%
  filter(gender == "Female") %>%
  distinct(cohort, .keep_all = TRUE) %>%
  mutate(percentage = round(women_cohort_count / cohort_count * 100, 2)) %>%
  select(cohort, cohort_count, women_cohort_count, percentage) %>%
  arrange(desc(percentage))

kable(gender_cohort)


```

```{r}
#same last name and same corporation 
family_corporations = all_doctors_raw %>%
  select(name, cpso, active_member, educ_year, gender, specialties_1) %>%
  drop_na(name, cpso, active_member) %>%
  mutate(last_name = str_extract(name, "^[^,]+")) %>%
  select(-active_member)

family_long_corp = long_corp_two %>%
  group_by(corporation) %>%
  mutate(owner_count = n())

family_corporations = left_join(family_corporations, family_long_corp, by =  c("cpso"))

family_corporations <- family_corporations %>%
  drop_na(corporation) %>%
  group_by(last_name, corporation) %>%
  mutate(family_corp = ifelse(
    n_distinct(cpso) > 1 & 
    owner_count == 2 & 
    any(gender == "Male") & 
    any(gender == "Female") & 
    (max(educ_year) - min(educ_year) <= 10),  # Education year gap condition
    1, 0
  )) %>%
  ungroup() %>%
  filter(family_corp == 1)

duplicates_df <- family_corporations %>%
  group_by(cpso) %>%
  filter(n() > 1) %>%  # Keep only names that appear more than once
  ungroup()
  

#here is to see the most common specialties for family corporations
family_specialties = family_corporations %>%
  group_by(specialties_1) %>%
  mutate(specialties_count = n()) %>%
  ungroup() %>%
  distinct(specialties_1, .keep_all = TRUE) %>% 
  select(specialties_1, specialties_count)

#see how many have the same specialty 
same_speciality <- family_corporations %>%
  filter(family_corp == 1) %>%
  group_by(corporation) %>%
  mutate(same_specialty = ifelse(n_distinct(specialties_1) == 1, 1, 0)) %>%
  ungroup()

same_speciality_prop = (sum(same_speciality$same_specialty == 1)/ nrow(same_speciality))

#here is to see all corporations that are considered family practices
family_corp_distinct = family_corporations %>%
  distinct(corporation, last_name, .keep_all = TRUE) %>%
  select(-c(cpso, specialties_1, educ_year, name, gender))

#here is to see all doctors that are in a family practice
family_doctors_distinct = family_corporations %>%
  distinct(cpso, .keep_all = TRUE)

#here is all corporations
total_corps = long_corp_two %>%
  distinct(corporation, .keep_all =  TRUE)

total_family_corps = sum(family_corp_distinct$family_corp == 1) #total family corporations
total_active_doctors = sum(!is.na(clean_all_doctors$active_member)) # total active doctors
total_family_doctors = sum(family_doctors_distinct$family_corp == 1) #total active doctors that are in family coprs
total_corps = nrow(total_corps) #total corporations

```

## Summary Stats on *Family Practices*

-   **Total Family Practices:** 166\
-   **Doctors in Family Practices:** 316\
-   **Percentage of Corporations that are Family Practices:** **0.80%**\
-   **Percentage of Active Doctors in a Family Practice:** **0.86%**

### **Top Specialties in Family Practices**

1.  **Family Medicine** – 139\
2.  **Internal Medicine** – 38\
3.  **Pediatrics** – 21\
4.  **Anesthesiology** – 17\
5.  **Psychiatry** – 17

**Family Practices where Both Doctors Share the Same Specialty:** **28.3%**
